---
title: "Week 2: Exploring TC journals"
---

## Install packages and read in data

Our source data is split between two CSV files, one for the journal "Technical Communication Quarterly" and one for "Technical Communication."

```{r}
# install and load packages

#install.packages("tidyverse")
library(tidyverse)
library(readr)

# read in both csv files
tcq_data_raw <- read_csv("data/tcq_wos_data.csv")
tc_data_raw <- read_csv("data/tc_wos_data.csv")
```

## Data preparation 

Let's take a quick look at the data and clean up any issues. Some questions we might answer:

* How many rows and columns are in each dataset?
* What do the rows and columns correspond to?

```{r}
# shows dimensions (number of rows and columns)
dim(tcq_data_raw) 
dim(tc_data_raw)

glimpse(tcq_data_raw)
glimpse(tc_data_raw)
```

### Cleaning names

We'll start by cleaning up the names with the janitor package. The convention is lowercase and joined by underscores. 

```{r}
#install.packages("janitor")
library(janitor)

# use help("library_name") for a description
help("janitor") 


# Syntax method 1 (on TCQ)
tcq_data_raw <- clean_names(tcq_data_raw)

glimpse(tcq_data_raw)


# Syntax method 2 (on TC)
tc_data_raw <- tc_data_raw %>%
  clean_names()

glimpse(tc_data_raw)

```

### Select columns of interest

Now we'll use a function from a library called dplyr to retain only the columns we want. Let's keep:

* author_full_names
* article_title
* source_title
* abstract
* cited_references
* cited_reference_count
* publication_year
* publication_type

First on TCQ

```{r}
library(dplyr)

# the select function from dplyr on TCQ data
tcq_clean <- tcq_data_raw %>%
  select(author_full_names,
         article_title,
         source_title,
         abstract,
         cited_references,
         cited_reference_count,
         publication_year,
         publication_type)

glimpse(tcq_clean)

```

Now on TC

```{r}
# the select function from dplyr on TCQ data
tc_clean <- tc_data_raw %>%
  select(author_full_names,
         article_title,
         source_title,
         abstract,
         cited_references,
         cited_reference_count,
         publication_year,
         publication_type)

glimpse(tc_clean)
```

### Combine the two sets

```{r}

both_data <- rbind(tcq_clean, tc_clean)

dim(both_data)
```

## Exploring the data

### Some quick ways to examine at a high level

```{r}
# provides overview of numeric variables
summary(both_data) 


# overview of target column
summary(both_data$publication_year)

# for each column, shows data types and first few observations
str(both_data) 

# shows first few rows
head(both_data)

# shows the first n rows

head(both_data, n = 10)

#shows last few rows
tail(both_data)

# creates a frequency table for a categorical variable 
table(both_data$publication_year) 

```

### Visualize

Let's create some exploratory visualizations. 

#### Visualize the articles published by year

```{r}
# count of articles by year
ggplot(both_data, aes(x = publication_year)) +
  geom_bar() +
  labs(title = "Articles per Year in TC and TCQ",
       X = "Publication Year",
       Y = "Number of Articles") +
  theme_light()
```

#### Add another variable to the display

```{r}
# count of articles by year, by journal
ggplot(both_data, aes(x = publication_year, fill = source_title)) +
  geom_bar() +
  labs(title = "Articles per Year in TC and TCQ",
       X = "Publication Year",
       Y = "Number of Articles") +
  theme_light()
```

#### Create two displays using facet wrap

```{r}
# count of articles by year, by journal
ggplot(both_data, aes(x = publication_year)) +
  geom_bar() +
  labs(title = "Articles per Year in TC and TCQ",
       X = "Publication Year",
       Y = "Number of Articles") +
   facet_wrap(~ source_title, ncol = 2) # Change ncol as needed
 
```

#### Spice things up with a theme from ggthemes

Check out this [ggthemes gallery](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)

```{r}
#install.packages("ggthemes")
library(ggthemes)

# count of articles by year, by journal
ggplot(both_data, aes(x = publication_year)) +
  geom_bar() +
  labs(title = "Articles per Year in TC and TCQ",
       X = "Publication Year",
       Y = "Number of Articles") +
   facet_wrap(~ source_title, ncol = 1) +
  theme_economist()
```


```{r}
# count of articles by year, by journal
year_plot <- ggplot(both_data, aes(x = publication_year)) +
  geom_bar() +
  labs(title = "Articles per Year in TC and TCQ",
       X = "Publication Year",
       Y = "Number of Articles") +
   facet_wrap(~ source_title, ncol = 1) +
  theme_economist_white()

year_plot
```

#### Add labels above the bars

```{r}
# add labels above the bars
year_plot <- year_plot +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 4)

year_plot
```

#### Save the chart

```{r}
help(ggsave)
ggsave("plots/articles_barplot.png", 
       plot = year_plot, 
       width = 8, 
       height = 6, 
       dpi = 300,
       bg = NULL)
```

#### Save our dataset

```{r}
# save as csv

write.csv(both_data, "data_out/both_data.csv")

save(both_data, file = "data_out/both_data.RData")

```

## What about other journals?

Gather and add to our visuals all articles published after 2005 from [Web of Science](https://www.webofscience.com/wos/woscc/basic-search) in these journals:

* Journal of Business and Technical Communication
* Written Communication

